ANNO: The Anthropological Notation Ontology

### 2023-03-30 paper.tex
More ontology creation.

\section{Related Work}
\subsection{Anatomical Nomenclatures}
There are various nomenclatures for the description of anatomical landmarks.
These allow a uniform classification of anatomical features and structures:

@@

\subsection{\aw}

\section{Ontology Modelling}
% @Alex ist das CSV->OWL Script irgendwo publiziert, kann man das zitieren?
The domain experts are provided a spreadsheet-based input template by the ontologists,
which is transformed to OWL using a custom script.
% ist das richtig übersetzt?
Sources are categorized into core, root, and occurrence sources.

@@

TODO: Modellierung der Anthropologie-Domäne (u.a. Motivation, related work, was ist neu).
TODO: Dann könnte man Design-Entscheidungen beschreiben (wie und warum wurden bestimmte Sachverhalte so in der Ontologie modelliert).
\section{Publication}
ANNO is published on \url{https://ols.imise.uni-leipzig.de} as well as \url{https://annosaxfdm.de/ontology/}.
TODO: Bereitstellung der Ontologie für die Community (OLS etc.). 
\section{Integration into \aw{}}
TODO: Integration der Ontologie in die Anwendung.
%\subsection{}\label{s1.1}

>>>>

\section{Related Work}
There are various nomenclatures for the description of anatomical landmarks.
These allow a uniform classification of anatomical features and structures:

@@

\section{Ontology creation}

% take care to prevent the problems noted by reviewer 1 in https://semantic-web-journal.net/content/modelling-digital-health-data-examode-ontology-histopathology
% TODO: ontology language, expressivity
% TODO: reuse of ontological entities
% TODO: clear validation mechanism
% TODO: top-level ontology?
% @Alex ist das CSV->OWL Script irgendwo publiziert, kann man das zitieren?
The domain experts are provided a spreadsheet-based SMOG~\citep{smog} template by the ontologists, which is transformed to an OWL 2 ontology.
% ist das richtig übersetzt?
Sources are categorized into core, root, and occurrence sources.

% Aus dem Projektbericht übersetzt
\subsection{Selection criteria}
Initially, selection criteria were made for the development of the ontology.
All bones were to be integrated.
For the individual anatomical landmarks, it was necessary to take all those that occur in the TA with the exception of anatomical variants.
However, for the cranium, the selection was reduced because the number of all ossa cranii is too high.
Overall, however, it should include those that are of anthropological relevance, i.e., contribute to navigation, localization, and identification on the bone.
Those landmarks included in the definitions of others were also to be defined.
For bones that lie on the median (e.g., mandible or sternum), bilateral landmarks were defined with side reference in each case.
For osteometric landmarks, all those already established in the core literature were to be taken.
Furthermore, those were to be taken that are relevant for meaningful measurement distances and can be annotated in \aw{}.
For measurement distances, those that are established in the bulk of the literature and core literature as well as necessary for discriminant functions and functions for body height and body weight estimation should be selected.
In addition, it also had to be integrable into \aw{}.
For the functions, those with diagnostic value were taken.
These were discriminant functions for sex determination and regression functions for body height and body weight estimation.

\subsection{Process for creating definitions and measurements}
For the definitions and measurement, a minimum amount of English-language and mostly German-language literature was compared in order to develop the definitions from their information.
Notably, Latin or ancient Greek terms were missing from the English-language literature.
The FMA also contains only English terms.
For this reason, these were also searched out with their synonyms.
Overall, the name of the landmark or survey route had to be noted in English, German, and Latin in singular and plural form, its synonyms in the three languages, the FMA and TA ID, and any information on function and delineation.
Where there were contradictions in the literature, the information was highlighted, reconciled, and logically checked.
For the osteometric landmarks, the abbreviations and their synonyms of the three languages were also provided.
Furthermore, for these as well as for the measured sections, the sources were divided into parent source, core source and occurrence source.
The parent source designates the original source of the landmark.
Under the core source, those were listed where the landmarks are listed and sampled by default.
Occurrence source represents the sources where the landmark occurs for measurements.
For redefinitions of landmarks, a meaningful Latin name had to be selected.
Requirements for this were an included position description (e.g. Punctum superioris capitis femoris as superior located point of the Caput femoris) necessary.
For this, the examination of the Latin grammar was relevant.
For the measurement sections, the type of measurement (e.g., distance measurement) and the measurement instrument were taken in addition to the name of the measurement section.
The subsequently created visual definitions were made in the different anatomical views and represented area markings for the anatomical landmarks and point markings for the osteometric ones.

\subsection{Functions}
The functions were divided into discriminant functions for sex determination and regress functions for body height and body weight estimation.
The goal here was to cover functions for at least one European, African, American, and Asian ethnicity or population.
Names were assigned by the number included in each study, the authors, and the year of publication.
In addition, the function, reference population, aspect (e.g., discriminant function), and sample size with division by gender were noted.
In addition, for the discriminant functions, the thresholds of sex assignment, classification accuracy, and misclassification were important; for the regression functions, the error interval was important.

@@

TODO: Modellierung der Anthropologie-Domäne (u.a. Motivation, related work, was ist neu).
TODO: Dann könnte man Design-Entscheidungen beschreiben (wie und warum wurden bestimmte Sachverhalte so in der Ontologie modelliert).
\section{Publication}
ANNO is published on \url{https://ols.imise.uni-leipzig.de} as well as \url{https://annosaxfdm.de/ontology/}.
TODO: Bereitstellung der Ontologie für die Community (OLS etc.).

\section{Use Case: Integration into \aw{}}
% this is based on project report but has to be rewritten: less about AW3d and more about the usage of ANNO by it
ANNO is used by \aw{}~\citep{anthroworks3d}.
By combining user-friendly techniques of photogrammetry, insights from user experience research and knowledge from game development, a digital twin is created, which can subsequently be examined virtually.
This enables location-independent and parallel work without wear and tear on the bone.
The examination can be performed as often as desired, even if the skeletal individuals or collections are not available at the institute or have already been reburied.
It also draws its advantage when space is limited.
Thus, it requires only three SLR cameras, sufficient exposure and a computer.
The examination proceeds as follows: After photogrammetry is done and the 3D model is created, the objects are imported into the software and calibrated.
For calibration, there is an individual placeholder for each bone, which must first be selected.
Then the calibrated digital twin is placed in a placeholder skeleton.
Annotations can be made in the detailed view of each bone.
Point, line and area markings are available for selection.
Furthermore, it is possible to make measurements.
You can choose between distance, angle and circumference measurements.
After marking or measurement, an identification number is assigned to each one.
There is also a documentation about the time of the annotation as well as the name of the editor.
The marking or measurement can be classified according to certain categories (e.g. anatomical variation).
In addition, annotations are possible in a text field.
The annotations can be edited at any time afterwards, and the date of editing is recorded together with the author.
The annotations are subsequently displayed in different colors, so that overlaps can be kept apart.
A special feature of the software is the automation of measurements.
The program automatically sets pins for osteometric landmarks in a template view.
The person using the program can move these as desired.
The pins displayed have different color shades.
These differ depending on the relevance.
If the osteometric landmark is present in many measurement sections, its relevance increases and the pin appears darker.
After the pins have been roughly set, they can be refined using various views.
In the views, the individual osteometric landmarks are visible from different views, so that a fine adjustment can be made.
This is followed by the automatic measurement.
The individual values for each measurement section can be downloaded as a CSV file.
In addition, it is now possible to perform sex determinations using discriminant functions.
For this purpose, the required measurement sections can be selected.
The results are only visible in one display.
The automation allows a time saving, whereby more bones or skeletons can be examined in a shorter time.

%\subsection{}\label{s1.1}
